---
title: "Introduction to SQAPI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SQAPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is intended to guide users through the whole process of using SQAPI for querying the SQUIDLE+ API.
The intention is to introduce users to the SQAPI functions and provide useful workflow examples of how the package is can be used.
The most common use for SQAPI is to make GET requests. These send a http request to a specified endpoint with attached filters and parameters to get your data in a specified format.
This package also supports making POST and PATCH requests. 

For specific details on endpoints, see SQUIDLE+ API documentation: https://squidle.org/api/help?template=api_help_page.html

```{r setup}
library(SQAPI)
```
Upon loading the package, a start-up message informs Windows users of potential timeout issues with large HTTP requests. Follow these instructions if needed.

FLOWCHART HERE TO SHOW WORKFLOW


# Set up
To make requests to SQUIDLE+ API, you will first need to create an instance of SQAPI. This stores your token and the host URL. 
```{r, include=TRUE, eval=FALSE}
# Step 1: Create an instance of SQAPI
api <- SQAPI$new()
```
There is no need to actually pass the auth to this SQAPI$new() as it will interactively prompt you to enter your token. It will then store it to be passed to SQAPI::request() and SQAPI::export() functions. For security reasons, it is best to use this interactive feature and not hard code your token.
You will need to activate your token at https://squidle.org/.

# A GET request example


 
